Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

### 2024-05-02 docs/_tex/index.tex
adding bede's comments

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={PREPRINT -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- PREPRINT.},
  pdfauthor={Simon Oiry; Bede Ffinian Rowe Davies; Ana I. Sousa; Philippe Rosa; Maria Laura Zoffoli; Guillaume Brunier; Pierre Gernez; Laurent Barillé},
  pdfkeywords={Drone, Remote Sensing, Seagrass, Coastal
Ecosystems, Neural Network},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{PREPRINT -- Discriminating Seagrasses From Green Macroalgae in
European Intertidal areas using high resolution multispectral drone
imagery -- PREPRINT.}
\author[1]{Simon Oiry%
\corref{cor1}%
}
 \ead{oirysimon@gmail.com} 
\author[1]{Bede Ffinian Rowe Davies%
%
}

@@

\begin{abstract}
Coastal areas support seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
seagrass decline and habitat fragmentation. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the use of satellite data within monitoring programs. Yet,
satellite remote sensing struggles with spatial and spectral resolution,
making it difficult to discriminate seagrass from other macrophytes in
highly heterogenous meadows. Drone images at a very high spatial
resolution offer a promising solution to address challenges related to
spatial heterogeneity and intrapixel mixture. This study focuses on
using drone acquisitions with a ten spectral band sensor mirroring those
of Sentinel-2, for mapping intertidal macrophytes and effectively
discriminating between seagrass and green macroalgae. Nine drone flights
were conducted at two different altitudes (12m and 120m) across
heterogeneous intertidal European habitats in France and Portugal. Low
altitude flights were used to train a Deep Learning classifier based on
Neural Networks to discrimintate among five taxonomic classes of
intertidal vegetation: Magnoliopsida (Seagrass), Chlorophyceae (Green
macroalgae), Phaeophyceae (Brown algae), Rhodophyceae (Red macroalgae)
and benthic Bacillariophyceae (Diatoms). Classification of drone imagery
resulted in an overall accuracy of 94\% across all the sites and images,
covering a total area of 467 000 m². The model exhibited an accuracy of
96.4\% in identifying seagrass. Importantly, seagrass and green algae
can be discriminated, although they share the same pigment composition.
As the algorithm was developed for a multispectral camera with ten
spectral bands in the visible and near-infrared, it could be adapted to
the Multi-Spectral Instrument (MSI) onboard Sentinel-2 thus offering
promising perspectives for satellite remote sensing of intertidal
biodiversity over lager scales.
\end{abstract}

@@

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between the land and oceans \citep{unsworth2022}. Seagrass meadows
provide a myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production, protection against sea-level rise and
coastline erosion, and limitation of eutrophication. They serve as vital
habitats for a diverse array of marine and terrestrial species,
providing living, breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation are currently observed due to
climate change, diseases, urbanization, land reclamation, dredging,
competition with alien species, and reduction in water quality
\citetext{\citealp[ ]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp[ ]{duffy2019}; \citealp[
]{rasheed2011long}; \citealp{chefaoui2018dramatic}}. Both habitat
fragmentation and reduction, in turn, can severely copromise the
effectiveness of ecosystem services provided by seagrass meadows. While
improvements in water quality have been recently reported in Europe,
allowing an overall recovery of seagrass ecosystems at local scale, many
coastal waters worldwide are still subjected to strong eutrophication
processes \citetext{\citealp[ ]{deSantos2019}; \citealp{Zoffoli2021}}.
Coastal eutrophication has been ass ociated to excessive accumulation of
green macroalgae, so-called green tides (\citep{devlin2023nutrients}).
Green tides produce shade and suffocation over seagrass individuals,
thus threatening the health of seagrass ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs to monitor Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Traditionally, indicators of seagrass status have been quantified using
\emph{in situ} measurements. The acquisition of field data in intertidal
zones is however notoriously challenging. Intertidal seagrass meadows
are only partially exposed during low tide and can be situated in
difficult-to-reach mudflats, potentially leading to inaccurate and
limited estimations with conventional sampling techniques
\citep{nijland2019}. Satellite observations have been proven effective
in complementing in situ sampling, allowing for the near real-time and
consistent retrieval of seagrass EOVs and EBVs over extensive meadows.
\citetext{\citealp[ ]{Zoffoli2021}; \citealp[ ]{xu2021}; \citealp[
]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing (RS) provides temporally consistent
observations over large spatial scales, its utilization over intertidal
areas is limited by several constraints. Satellite missions with a high
temporal resolution (e.g.~daily MODIS observation) are limited by a too
coarse spatial resolution (\textgreater100m) to accurately map patchy
seagrass meadows. Missions with a high spatial resolution such as
Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low spectral
resolution. The limited number of spectral bands is indeed a challenge
to accurately discriminate seagrass from others co-existing macrophytes.
In particular Chlorophyceae (green algae) and marine Magnoliopsida
(seagrass) share the same pigment composition \citetext{\citealp[
]{ralph2002}; \citealp{Douay2022}}. Therefore, to someone not
specialized in the field, their spectral signatures may appear to be
alike \citetext{\citealp[ ]{Davies2023}; \citealp{bannari2022}}.
Recently, using advanced machine-learning algorithms trained with a
large hyperspectral library of more than 300 field reflectance spectra,
\citep{Davies2023} demonstrated that it was possible to discriminate
Magnoliopsida from Chlorophyceae using radiometric data acquired at
Sentinel-2 's spectral resolution. However the application of this
approach to satellite RS remains to be validated. Moreover patches of
green algae can develop at small spatial scales that are not observable
using non-commercial satellite imagery \citep{tuya2013}, especially
during the initial stage of a green tide.

Drones can potentially fill the gaps left by satellite RS and in situ
measurements, due to their ability to provide spatially-explicit
observations at very high spatial resolutions (pixel size from mm to cm)
while capturing data at multi-spectral resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping species
distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, most case studies were previously limited to
a single flight, restricting the generalizability of their application
over wider geographical scales \citetext{\citealp[
]{Roman2021}; \citealp[ ]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone RS to map
intertidal macrophytes over a diverse biogeographical range, with a
particular focus on discriminating Magnoliopsida and Chlorophyceae
(Seagrass and Green Algae respectively). Ten drone flights were
performed over soft-bottom intertidal areas along two European countries
(France and Portugal), covering a wide range of habitats, from
monospecific seagrass meadows to meadows mixed with green, or red
macroalgae. A deep learning algorithm was trained and validated for
macrophyte discrimination, emphasizing applicability across diverse
sites without a loss of prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in the Gulf of Morbihan (Figure~\ref{fig-map} A), France
(47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only
connected to the sea through a 900m wide channel. A total of 53 small
islands are scattered across the gulf leading to 250 km of shorelines.
Patchy seagrass meadows can be found on a lot of these islands. Within
the Gulf of Morbihan we chose two sites : one located on one of the
gulf's islands (Arz island) and the other located in the southern part
of the gulf (Duer). Two other sites were located in Bourgneuf Bay,
France (46.9849°N, 2.1488°W). This bay is 340 km² semi-enclosed
macrotidal bay, protected from waves by Noirmoutier Island. Bourgneuf
bay hosts a large intertidal seagrass meadow of about 6 km². Within this
meadow, the sites observed by drones (L'Epine and Barbatre,
Figure~\ref{fig-map} B) contained monospecific beds of \emph{Nanozostera
noltei} (dwarf eelgrass) with very little mixing with other macrophytes.
Three sites were surveyed in the Ria de Aveiro Lagoon in Portugal
(40.6887°N, 8.6810°W). The extent of this lagoon is about 80 km² with
many narrow channels, large salt marshes and many mudflats that uncovers
at low tide. It is connected to the open sea through a single channel,
with a tidal lag between the North and the South of the lagoon. The
southernmost site (Gafanha) was a mudflat located in a tributary river
of the lagoon whereas the two other sites (Mataducos and Marinha
Lanzarote) were situated in the middle of the lagoon and only accessible
by boat (Figure~\ref{fig-map} C). These Portuguese sites are
characterized by a more diverse intertidal vegetation, where patches of
seagrass intermingle with red, brown and green macroalgae.

@@

\caption{\label{tbl-flights}List of drone flight, summarising the date,
the altitude and the purpose of each flight. 12 m and 120 m flights have
a spatial resolution of 8 and 80 mm respectively.}

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25 m² assessing the presence or absence of five key taxonomic
classes of intertidal vegetation : Bacillariophyceae (Unicellular
benthic diatoms forming biofilms at the sediment surface during low
tide), Phaeophyceae (brown macroalgae), Magnoliopsida (dwarf seagrass),
Chlorophyceae (green macroalgae) and Rhodophyceae (red macroalgae)
(Figure~\ref{fig-vegetation}). Pictures of each quadrat were uploaded
online to the Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}, a public open portal to store and share biodiversity
data. Each photograph was also processed to estimate the percent cover
of each type of vegetation using an image processing software (imageJ).
For all quadrats, the hyperspectral reflectance signatures of each
vegetation class was measured using an ASD FieldSpec HandHeld 2
spectroradiometer, which acquires reflectance between 325 and 1075 nm,
with 1 nm of spectral resolution. Hyperspectral signatures from quadrats
serve for dual purposes: they validate the radiometric calibration of
drone data and contribute to error reduction in photo interpretation.
The workflow developed in this study is presented in
Figure~\ref{fig-workflow}.

@@

A structure-from-motion photogrammetry software (Agisoft Metashape) was
used to process images to obtain multispectral orthomosaics of each
flight. The workflow for orthomosaicking was the same for every flight.
First, tying key points were detected inside of each image and between
overlapping images in order to obtain a sparse point cloud. This cloud
was cleaned using reprojection accuracy metric in order to remove noisy
points. A dense point cloud was then produced using a structure from
motion algorithm. A surface interpolation of this dense point cloud was
made to obtain a digital surface model (DSM), used to reconstruct the
multispectral ortho-image. Low altitude drone flights produced
ortho-images with a very high spatial resolution (8 mm per pixel),
making it efficient to visually distinguish between the various types of
vegetation. High altitude fights on the other hand allowed to cover
large areas and produced images with a pixel size of 80 mm
(Table~\ref{tbl-flights}).

@@

\caption{\label{fig-workflow}Schematic representation of the workflow.
Diamonds represent input or output data, and rectangles represent Python
processing algorithms. The overall workflow of this study is divided
into two distinct parts based on the spatial resolution of the drone
flights: high-resolution flights (pixel size: 8 mm) were utilized for
training and prediction of the Neural Network model, whereas
lower-resolution flights (pixel size: 80 mm) were solely employed for
prediction purposes. Validation has been performed on both high and low
resolution flights.}

@@

\subsubsection{Neural Network model
building}\label{neural-network-model-building}

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model with 2 hidden layers. The training
pixels were categorized into seven different classes, representing the
various habitats encountered at the different study sites: Sediment,
Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table~\ref{tbl-flights}) were
used for training because their 8 mm spatial resolution allowied to
avoid spectral sub-pixel mixing and to accurately identify vegetation
classes. More than 418,000 pixels at 8 mm resolution from the 3 training
flights were used to train the model (Table~\ref{tbl-validationPX}).
Twenty one variables were used by the model as predictors: the ten raw
spectral bands of the Micasense RedEdge Dual MX multispectral camera
(ranging from 444 nm to 840 nm), the same ten spectral bands
standardized using a min/max transformation (Equation~\ref{eq-std} ;
\citep{Cao2017}) and the Normalized difference vegetation index (NDVI,
Equation~\ref{eq-ndvi}). Standardisation of spectral bands is commonly
used to eliminate the scaling differences between spectra and to limit
the effect of biomass on the spectra shape \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

@@

Variable Importance Plots (VIP) serve as a method to identify which
predictors are important for predicting a specific class. This is
achieved by repeatedly predicting the same dataset while shuffling one
predictor at a time. The benchmark score obtained after each iteration
is then compared to the benchmark score obtained without shuffling any
variables. The greater the difference between these two benchmark
values, the more important the variable is for the model
\citep{WEI2015399}.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft},
  pdfauthor={Simon Oiry; Bede Ffinian Rowe Davies; Ana I. Sousa; Philippe Rosa; Maria Laura Zoffoli; Guillaume Brunier; Pierre Gernez; Laurent Barillé},
  pdfkeywords={Drone, Remote Sensing, Seagrass, Coastal
Ecosystems, Neural Network},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Draft -- Discriminating Seagrasses From Green Macroalgae in
European Intertidal areas using high resolution multispectral drone
imagery -- Draft}
\author[1]{Simon Oiry%
\corref{cor1}%
}
 \ead{oirysimon@gmail.com} 
\author[1]{Bede Ffinian Rowe Davies%
%
}

@@

\begin{abstract}
Coastal areas support seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
habitat fragmentation and seagrass decline. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the addition of satellite data within monitoring programs.
Yet, satellite remote sensing struggles with spatial and spectral
resolution, making it difficult to discriminate seagrass from other
macrophytes in highly heterogenous meadows. Drone images at a very high
spatial resolution offer a promising solution to address challenges
related to spatial heterogeneity and intrapixel mixture. This study
focuses on using drone acquisitions with a ten spectral band sensor
mirroring those of Sentinel-2, for mapping intertidal macrophytes and
effectively discriminating between seagrass and green macroalgae. Nine
drone flights were conducted at two different altitudes (12m and 120m)
across heterogeneous intertidal European habitats in France and
Portugal. Low altitude flights were used to train a Deep Learning
classifier based on Neural Networks to discrimintate among five
taxonomic classes of intertidal vegetation: Magnoliopsida (Seagrass),
Chlorophyceae (Green macroalgae), Phaeophyceae (Brown algae),
Rhodophyceae (Red macroalgae) and benthic Bacillariophyceae (Diatoms).
Classification of drone imagery resulted in an overall accuracy of 94\%
across all the sites and images, covering a total area of 467 000 m².
The model exhibited an accuracy of 96.4\% in identifying seagrass.
Importantly, seagrass and green algae can be discriminated, although
they share the same pigment composition. As the algorithm was developed
for a multispectral camera with ten spectral bands in the visible and
near-infrared, it could be adapted to the Multi-Spectral Instrument
(MSI) onboard Sentinel-2 thus offering promising perspectives for
satellite remote sensing of intertidal biodiversity over lager scales.
\end{abstract}

@@

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between the land and oceans \citep{unsworth2022}. Seagrass meadows
provide a myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production, protection against sea-level rise and
coastline erosion, and limitation of eutrophication. They serve as vital
habitats for a diverse array of marine and terrestrial species,
providing living, breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation of seagrass meadows are currently
observed due to climate change, diseases, urbanization, land
reclamation, dredging, competition with alien species, and reduction in
water quality \citetext{\citealp[ ]{nguyen2021}; \citealp[
]{soissons2018}; \citealp[ ]{orth2006}; \citealp[ ]{lin2018}; \citealp[
]{duffy2019}; \citealp[
]{rasheed2011long}; \citealp{chefaoui2018dramatic}}. Both habitat
fragmentation and reduction, in turn, can severely compromise the
effectiveness of ecosystem services provided by seagrass meadows. While
improvements in water quality have been recently reported in Europe,
allowing an overall recovery of seagrass ecosystems at local scales,
many coastal waters worldwide are still subjected to strong
eutrophication processes \citetext{\citealp[
]{deSantos2019}; \citealp{Zoffoli2021}}. Coastal eutrophication has been
associated to excessive accumulation of green macroalgae, so-called
green tides \citep{devlin2023nutrients}. Green tides produce shade and
suffocation over seagrass individuals, thus threatening the health of
seagrass ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs to monitor Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Traditionally, indicators of seagrass status have been quantified using
\emph{in situ} measurements. However, the acquisition of these
measurements in intertidal zones is notoriously challenging. Intertidal
seagrass meadows are only partially exposed during low tide and can be
situated in difficult-to-reach mudflats, potentially leading to
inaccurate and limited estimations with conventional sampling techniques
\citep{nijland2019}. Satellite observations have been proven effective
in complementing \emph{in situ} sampling, allowing for the near
real-time and consistent retrieval of seagrass EOVs and EBVs over
extensive meadows. \citetext{\citealp[ ]{Zoffoli2021}; \citealp[
]{xu2021}; \citealp[ ]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing (RS) provides temporally consistent
observations over large spatial scales, its utilization over intertidal
areas is limited by several constraints. Satellite missions with a high
temporal resolution (e.g.~daily MODIS observation) are limited by a too
coarse spatial resolution (\textgreater100m) to accurately map patchy
seagrass meadows. Missions with a high spatial resolution such as
Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low spectral
resolution. The limited number of spectral bands challenges accurate
discrimination of seagrass from other co-existing macrophytes. In
particular Chlorophyceae (green algae) and marine Magnoliopsida
(seagrass) share the same pigment composition \citetext{\citealp[
]{ralph2002}; \citealp{Douay2022}}. Therefore, to someone not
specialized in the field, their spectral signatures may appear to be
alike \citetext{\citealp[ ]{Davies2023}; \citealp{bannari2022}}.
Recently, using advanced machine-learning algorithms trained with a
large hyperspectral library of more than 300 field reflectance spectra,
\citep{Davies2023} demonstrated that it was possible to discriminate
Magnoliopsida from Chlorophyceae using radiometric data acquired at
Sentinel-2 's spectral resolution. However the application of this
approach to satellite RS remains to be validated. Moreover patches of
green algae can develop at small spatial scales that are not observable
using non-commercial satellite imagery \citep{tuya2013}, especially
during the initial stage of a green tide.

Drones can potentially fill the data gaps left by satellite RS and
\emph{in situ} measurements, due to their ability to provide
spatially-explicit observations at very high spatial resolutions (pixel
size from mm to cm) while capturing data at multi-spectral resolution
\citetext{\citealp[ ]{fairley2022drone}; \citealp{oh2017use}}. The
versatility of drones allows for their application across a diverse
thematic range , from coastal zone management \citetext{\citealp[
]{adade2021}; \citealp[ ]{casella2020}; \citealp{angnuureng2022}} to
mapping species distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, most case studies were previously limited to
a single flight, restricting the generalizability of their application
over wider geographical scales \citetext{\citealp[
]{Roman2021}; \citealp[ ]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone RS to map
intertidal macrophytes over a diverse biogeographical range, with a
particular focus on discriminating Magnoliopsida and Chlorophyceae
(Seagrass and Green Algae respectively). Ten drone flights were
performed over soft-bottom intertidal areas along the Atlantic
coastlines of two European countries (France and Portugal), covering a
wide range of habitats, from monospecific seagrass meadows to meadows
mixed with green, or red macroalgae. A deep learning algorithm was
trained and validated for macrophyte discrimination, emphasizing
applicability across diverse sites without a loss of prediction
accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites were
located in the Gulf of Morbihan (Figure~\ref{fig-map} A), France
(47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only
connected to the sea through a 900m wide channel. A total of 53 small
islands are scattered across the gulf leading to 250 km of shorelines.
Patchy seagrass meadows can be found on many of these islands. One of
the sites within the gulf was on one its islands (Arz) and the other was
located further south on a mainland beach area (Duer). Two other sites
were located in Bourgneuf Bay, France (46.9849°N, 2.1488°W). This is a
340 km² semi-enclosed macrotidal bay, protected from waves by
Noirmoutier Island. Bourgneuf bay hosts a large intertidal seagrass
meadow of about 6 km² \citep{ZOFFOLI2020112020}. Within this meadow, the
sites observed by drones (L'Epine and Barbatre, Figure~\ref{fig-map} B)
contained monospecific beds of \emph{Nanozostera noltei} (dwarf
eelgrass) with very little mixing with other macrophytes. Three sites
were surveyed in the Ria de Aveiro Lagoon in Portugal (40.6887°N,
8.6810°W). The extent of this lagoon is \textasciitilde80 km² with many
narrow channels, large salt marshes and many mudflats that uncover at
low tide. It is connected to the open sea through a single channel, with
a tidal lag between the North and the South of the lagoon. The
southernmost site (Gafanha) was a mudflat located in a tributary river
of the lagoon whereas the two other sites (Mataducos and Marinha
Lanzarote) were situated in the middle of the lagoon and only accessible
by boat (Figure~\ref{fig-map} C). These Portuguese sites are
characterized by a more diverse intertidal vegetation, where patches of
seagrass intermingle with red, brown, and green macroalgae.

@@

\caption{\label{tbl-flights}List of drone flight, summarising the date,
the altitude, and the purpose of each flight. 12 m and 120 m flights
have a spatial resolution of 8 and 80 mm respectively.}

@@

Before each flight, targets used as ground control points were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). Ground control points were used to
correct georeferencing imprecision of orthomosaics with an horizontal
and vertical accuracy of 10cm. A dGPS was also used to georeference
quadrats of 0.25 m² ,which assessed the presence or absence of five key
taxonomic classes of intertidal vegetation : Bacillariophyceae
(Unicellular benthic diatoms forming biofilms at the sediment surface
during low tide), Phaeophyceae (brown macroalgae), Magnoliopsida (dwarf
seagrass), Chlorophyceae (green macroalgae) and Rhodophyceae (red
macroalgae) (Figure~\ref{fig-vegetation}). Pictures of each quadrat were
uploaded online to the Global Biodiversity Information Facility (GBIF)
platform \citep{BedeGbif}, a public open portal to store and share
biodiversity data. Each photograph was also processed to estimate the
percent cover of each type of vegetation using an image processing
software \citep[ImageJ]{schneider2012nih}. Hyperspectral reflectance
signatures of each vegetation class were recorded using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures
served dual purposes: they validate the radiometric calibration of drone
data and contribute to error reduction in photo interpretation.

@@

A structure-from-motion photogrammetry software (Agisoft Metashape) was
used to process images to obtain multispectral orthomosaics of each
flight. The process for orthomosaicking was identical for every flight.
First, tying key points were detected inside of each image and between
overlapping images in order to obtain a sparse point cloud. This cloud
was cleaned using reprojection accuracy metric in order to remove noisy
points. A dense point cloud was then produced using a structure from
motion algorithm. A surface interpolation of this dense point cloud was
made to obtain a digital surface model (DSM), used to reconstruct the
multispectral ortho-image \citep{nebel2020review}. Low altitude drone
flights produced ortho-images with a very high spatial resolution (8 mm
per pixel), making it efficient to visually distinguish between the
various types of vegetation. High altitude fights on the other hand
allowed to cover large areas and produced images with a pixel size of 80
mm (Table~\ref{tbl-flights}).

@@

\caption{\label{fig-workflow}Schematic representation of the workflow.
Parallelograms represent input or output data, and rectangles represent
Python processing algorithms. The overall workflow of this study is
divided into two distinct parts based on the spatial resolution of the
drone flights: high-resolution flights (pixel size: 8 mm) were utilized
for training and prediction of the Neural Network model, whereas
lower-resolution flights (pixel size: 80 mm) were solely employed for
prediction purposes. Validation has been performed on both high and low
resolution flights.}

@@

\subsubsection{Training dataset
building}\label{training-dataset-building}

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model. The training pixels were
categorized into seven different classes, representing the various
habitats encountered at the different study sites: Sediment, Water,
Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table~\ref{tbl-flights}) were
used for training because their 8 mm spatial resolution allowied to
avoid spectral sub-pixel mixing and to accurately identify vegetation
classes. More than 418,000 pixels at 8 mm resolution from the 3 training
flights were used to train the model (Table~\ref{tbl-validationPX}).
Twenty one variables were used by the model as predictors: the ten raw
spectral bands of the Micasense RedEdge Dual MX multispectral camera
(ranging from 444 nm to 840 nm), the same ten spectral bands
standardized using a min/max transformation (Equation~\ref{eq-std} ;
\citep{Cao2017}) and the Normalized difference vegetation index (NDVI,
Equation~\ref{eq-ndvi}). Standardisation of spectral bands is commonly
used to eliminate the scaling differences between spectra and to limit
the effect of biomass on the spectra shape \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

@@

Variable Importance Plots (VIP) serve as a method to identify which
predictors are important for predicting a specific class. Out of the
twenty one predictors utilized in this study, Variable Importance in
Projection (VIP) was computed only for the raw and standardized values
of the 10 spectral bands captured by the MicaSense camera. This is
achieved by repeatedly predicting the same dataset while shuffling one
predictor at a time. The benchmark score obtained after each iteration
is then compared to the benchmark score obtained without shuffling any
variables. The greater the difference between these two benchmark
values, the more important the variable is for the model
\citep{WEI2015399}.
